---
title: "DAPS Air Pollution Monitor Differences, Correlation and Basic Plots \n OUTLIERS REMOVED"
author: "Melissa Lowe"
date: "6/24/19"
output: pdf_document
---

#### In this case, NO OUTLIERS WERE REMOVED 

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.path='CorrPlots_N0_OUTLIERS/')
```


```{r, include=FALSE}
library(lme4)
library(lmtest)
library(dplyr)
library(readxl)
library(utils)
library(knitr)
library(ggplot2)
library(plyr)
library(lmerTest)
library(reshape2)
library(stargazer)
library(rmcorr)
library(RColorBrewer)

```


```{r, include=FALSE}
setwd("~/Desktop")
dapss <- read.csv("~/Desktop/DAPS_REPO/ReproducibleItemsforDAPS/DATA_PROCESSED/DAPS_2Cohort.csv")
daps <-dapss[, -1] #
```


# Basic Correlations

### Correlation of CDPHE Mean Ozone and Wearable Ozone
```{r, echo=FALSE}

#correlation of ozone
o3corr <- cor.test(daps$vest_O3, daps$CDPHE_O3_mean, method="pearson")
o3corr
```
### Plot
```{r, echo=FALSE}
o3corr_2 <- lm(CDPHE_O3_mean ~ vest_O3, data=daps)

plot(daps$vest_O3, daps$CDPHE_O3_mean, col="blue", lty= 3, pch=19, xlab=expression(paste("Wearable" , ~ O[3])), ylab = expression(paste("CDPHE" , ~ O[3])), main= expression(paste("Correlations between CDPHE and Wearable" , ~ O[3])))
abline(o3corr_2, col="black", lty=2, lwd=3)

text(12, 35, expression(bold(rho)[O[3]] == 0.075), cex=1)
text(12, 32, expression(italic(p-value) == 0.674), cex = 0.6)
```


### Correlation of Stationary PM1O and Wearable PM10

```{r, echo=FALSE}
#correlation of pm10
pm10corr <- rmcorr(id, log_vest_pm10, log_msp_pm10,daps)
pm10corr 
```

### Plot
```{r, echo=FALSE}

blueset <- brewer.pal(8, 'Blues')
pal <- colorRampPalette(blueset)
plot(pm10corr,  overall = TRUE, palette = pal, overall.col = 'black', lty = 3, xlab = expression(paste("Wearable" , ~ PM[10])), ylab = expression(paste("Stationary" , ~ PM[10])), main= expression(paste("Correlations between Stationary and Wearable" , ~ PM[10]))) 


text(5.25, 4, expression(bold(rho)[PM[10]] == 0.502), cex=1)
text(5.25, 3.75, expression(italic(p-value) == 5.13e-6), cex = 0.6)
```


### Correlation of CDPHE mean PM10 and wearable PM10
```{r, echo=FALSE}
pm10corr2 <-rmcorr(id, log_vest_pm10, log(daps$CDPHE_PM10_mean), daps)
pm10corr2
```

### Plot
```{r, echo=FALSE}
blueset <- brewer.pal(8, 'Blues')
pal <- colorRampPalette(blueset)
plot(pm10corr2, overall = TRUE, palette = pal, overall.col = 'black',lty = 3, xlab = expression(paste("Wearable" , ~ PM[10])), ylab = expression(paste("CDPHE" , ~ PM[10])), main= expression(paste("Correlations between CDPHE and Wearable" , ~ PM[10]))) 

text(5.25, 3.25, expression(bold(rho)[PM[10]] == -0.15), cex=1)
text(5.25, 3.15, expression(italic(p-value) == 0.21), cex = 0.6)
```

### Correlation of Stationary PM10 and CDPHE PM10
```{r, echo=FALSE}
pm10corr3 <- rmcorr(id, log_msp_pm10, log(CDPHE_PM10_mean), daps)
pm10corr3
```

### PLOT
```{r, echo=FALSE}

plot(pm10corr3, overall = TRUE, palette = pal, overall.col = 'black',lty = 3, xlab = expression(paste("Stationary" , ~ PM[10])), ylab = expression(paste("CDPHE" , ~ PM[10])), main= expression(paste("Correlations between CDPHE and Stationary" , ~ PM[10]))) 

text(5.2, 2.7, expression(bold(rho)[PM[10]] == -0.06), cex=1)
text(5.2, 2.6, expression(italic(p-value) == 0.59), cex = 0.6)



```

### Plot of Relative Correlation with Wearable Monitor for PM10 - includes log values
```{r, echo=FALSE}


b <- ggplot(daps)
b + 
geom_point(aes(log_vest_pm10, log_msp_pm10), color = "red1", shape =17)+
geom_smooth(aes(log_vest_pm10, log_msp_pm10), color = "red1", method="lm", linetype = "dotdash")+
geom_point(aes(log_vest_pm10, log(CDPHE_PM10_mean)),color= "mediumblue", shape = 19) +
geom_smooth(aes(log_vest_pm10, log(CDPHE_PM10_mean)), color = "mediumblue", method ="lm", linetype= "solid") +
annotate(geom="text", x=4.9, y=1.9, label="CDPHE vs. Wearable, \n corr = -0.15, p=0.21",
              color="mediumblue", fontface="bold")+
annotate(geom="text", x=5, y=5.2, label="Stationary vs. Wearable, \n corr = 0.5, p=5.1e-6",
              color="red1", fontface="bold")+
xlab(expression(paste("wearable" , ~ log(PM[10])))) +
ylab(expression(log(PM[10])))+ 
labs(title = "Relative Correlation with Wearable Monitor")+ 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
theme(plot.title=element_text( size=15)) +
theme(text = element_text(size = 15, color= "black"))

```


### Plot of Relative Correlation with Wearable Monitor for PM10 - includes unlogged values
```{r, echo=FALSE}

b <- ggplot(daps)
b + 
geom_point(aes(exp(log_vest_pm10), exp(log_msp_pm10)), color = "red1", shape =17)+
geom_smooth(aes(exp(log_vest_pm10), exp(log_msp_pm10)), color = "red1", method="lm", linetype = "dotdash")+
geom_point(aes(exp(log_vest_pm10), CDPHE_PM10_mean),color= "mediumblue", shape = 19) +
geom_smooth(aes(exp(log_vest_pm10), CDPHE_PM10_mean), color = "mediumblue", method ="lm", linetype= "solid") +
annotate(geom="text", x=200, y=40, label="CDPHE Mean vs. Wearable Monitor, \n corr = -0.15, p=0.21",
              color="mediumblue", fontface="bold")+
annotate(geom="text", x=150, y=150, label="Stationary vs. Wearable Monitor, \n corr = 0.5, p=5.1e-6",
              color="red1", fontface="bold")+
xlab(expression(paste("wearable" , ~ PM[10]))) +
ylab(expression(PM[10]))+ 
labs(title = "Relative Correlation with Wearable Monitor")+
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
theme(plot.title=element_text( size=15)) +
theme(text = element_text(size = 15, color= "black"))

```



### Brief look at health outcomes with exhaled nitric oxide and PM10.

Here, we can see the moderate weakness of the relationship between PM10 and FeNO.

```{r, echo=FALSE}

b <- ggplot(daps)
b + 
geom_point(aes(log_vest_pm10, log_FeNO_avg), color = "green3")+
geom_smooth(aes(log_vest_pm10, log_FeNO_avg), color = "green3")+
geom_point(aes(log_msp_pm10, log_FeNO_avg),color= "blue") +
geom_smooth(aes(log_msp_pm10, log_FeNO_avg), color = "blue") +
annotate(geom="text", x=4.5, y=1.5, label=expression(paste("log(FeNO) and Stationary" , ~ PM[10])),
              color="blue", size =4)+
  annotate(geom="text", x=2.5, y=5, label=expression(paste("log(FeNO) and Wearable" , ~ PM[10])),
              color="darkgreen", size=4)+
xlab(expression(log(PM[10]))) +
ylab("log(FeNO)")+ 
labs(title = expression(paste("Exhaled Nitric Oxide and", ~ PM[10] ~ "Exposure")))+ 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
theme(plot.title=element_text( face="bold", size=16)) +
theme(text = element_text(size = 16, color= "black")) 

```


### Correlation between stationary NO2 and wearable NO2
```{r, echo=FALSE}

#correlation of no2

no2corr <- rmcorr(id, log_vest_NO2, log_stat_NO2,daps)
no2corr
```

### Plot
```{r, echo=FALSE}
plot(no2corr, overall = TRUE, palette = pal, overall.col = 'black',lty = 3, xlab = expression(paste("Wearable" , ~ NO[2])), ylab = expression(paste("Stationary" , ~ NO[2])), main= expression(paste("Correlations between Stationary and Wearable" , ~ NO[2]))) 

text(3,1.4, expression(bold(rho)[NO[2]] == 0.95), cex=1)
text(3, 1, expression(italic(p-value) == 2.49e-36), cex = 0.6)
```


### Correlation between CDPHE mean NO2 and wearable NO2
```{r, echo=FALSE}

no2corr2 <- rmcorr(id,log_vest_NO2, log_CDPHE_NO2_mean, daps)
no2corr2
```

### Plot
```{r, echo=FALSE}
plot(no2corr2, overall = TRUE, palette = pal, overall.col = 'black',lty = 3, xlab = expression(paste("Wearable" , ~ NO[2])), ylab = expression(paste("CDPHE" , ~ NO[2])), main= expression(paste("Correlations between CDPHE and Wearable" , ~ NO[2]))) 

text(0,.95, expression(bold(rho)[NO[2]] == 0.23), cex=1)
text(0, .92, expression(italic(p-value) == 0.051), cex = 0.6)
```


### Correlation between CDPHE mean NO2 and stationary NO2
```{r, echo=FALSE}
no2corr3 <- rmcorr(id, log_stat_NO2, log_CDPHE_NO2_mean, daps)
no2corr3
```

### Plot
```{r, echo=FALSE}
plot(no2corr3, overall = TRUE, palette = pal, overall.col = 'black',lty = 3, xlab = expression(paste("Stationary" , ~ NO[2])), ylab = expression(paste("CDPHE" , ~ NO[2])), main= expression(paste("Correlations between CDPHE and Stationary" , ~ NO[2]))) 

text(0,.95, expression(bold(rho)[NO[2]] == 0.19), cex=1)
text(0, .92, expression(italic(p-value) == 0.098), cex = 0.6)

```


### Plot of Relative Correlation with Wearable Monitor for PM10 - includes logged values
```{r, echo=FALSE}
c <- ggplot(daps)
c + 
geom_point(aes(log_vest_NO2, log_stat_NO2), color = "red1", shape =17)+
geom_smooth(aes(log_vest_NO2, log_stat_NO2), color = "red1", method="lm", linetype = "dotdash")+
geom_point(aes(log_vest_NO2, log_CDPHE_NO2_mean),color= "mediumblue", shape = 19) +
geom_smooth(aes(log_vest_NO2, log_CDPHE_NO2_mean), color = "mediumblue", method ="lm", linetype= "solid") +
annotate(geom="text", x=0.25, y=4.5, label="CDPHE Mean vs. Wearable Monitor, \n corr = 0.23, p=0.05",
              color="mediumblue", fontface="bold")+
annotate(geom="text", x=2, y=0, label="Stationary vs. Wearable Monitor , \n corr = 0.95, p=2.5e-36",
              color="red1", fontface="bold")+
xlab(expression(paste("Wearable" , ~ log(NO[2])))) +
ylab(expression( log(NO[2])))+
labs(title = "Relative Correlation with Wearable Monitor")+ 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
theme(plot.title=element_text( face="bold", size=15)) +
theme(text = element_text(size = 15, color= "black")) 
```

### Plot of Relative Correlation with Wearable Monitor for PM10 - includes unlogged values
```{r, echo=FALSE}
c <- ggplot(daps)
c + 
geom_point(aes(exp(log_vest_NO2), exp(log_stat_NO2)), color = "red1", shape =17)+
geom_smooth(aes(exp(log_vest_NO2), exp(log_stat_NO2)), color = "red1", method="lm", linetype = "dotdash")+
geom_point(aes(exp(log_vest_NO2), exp(log_CDPHE_NO2_mean)),color= "mediumblue", shape = 19) +
geom_smooth(aes(exp(log_vest_NO2), exp(log_CDPHE_NO2_mean)), color = "mediumblue", method ="lm", linetype= "solid") +
annotate(geom="text",  x=30, y=0, label="CDPHE Mean vs. Wearable NO2, \n corr = 0.23, p=0.05",
              color="mediumblue", fontface="bold")+
annotate(geom="text", x=40, y=90, label="Stationary vs. Wearable NO2, \n corr = 0.95, p=2.5e-36",
              color="red1", fontface="bold")+
xlab(expression(paste("Wearable" , ~ NO[2]))) +
ylab(expression(NO[2]))+ 
labs(title = "Relative Correlation with Wearable Monitor")+ 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
theme(plot.title=element_text(face="bold", size=15)) +
theme(text = element_text(size = 15, color= "black"))

```


Correlation of glucans with endotoxins
```{r, echo = FALSE}


#correlation of glucans with endotoxins
glendocorr3 <- cor.test(daps$gluc_m3, log(daps$endo_m3), method="pearson")
glendocorr3

glendocorr2 <- lm(daps$gluc_m3 ~ log(daps$endo_m3), data=daps)

plot(log(daps$endo_m3),daps$gluc_m3 , col="blue", lty= 3, pch=19, xlab=("Log(Endotoxin) Value"), ylab = ("Glucan Value"), main= ("Correlations between Glucans and Endotoxins"))
abline(glendocorr2, col="black", lty=2, lwd=3)

text(-1, 400, expression(bold(rho) == 0.51), cex=1)
text(-1, 370, expression(italic(p-value) == 0.0006), cex = 0.7)


```


