---
title: "Models and Time-Varying Plots"
author: "Melissa Lowe"
date: "6/21/2019"
output: pdf_document
---

####  In this case, no outliers were removed.
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.path='TVMonitorPlots/')
```

```{r, include=FALSE}
library(lme4)
library(lmtest)
library(dplyr)
library(readxl)
library(utils)
library(knitr)
library(ggplot2)
library(plyr)
library(lmerTest)
library(reshape2)
library(stargazer)
library(rmcorr)
library(RColorBrewer)

```


```{r, include=FALSE}
setwd("~/Desktop")
dapss <- read.csv("~/Desktop/DAPS_REPO/ReproducibleItemsforDAPS/DATA_PROCESSED/DAPS_2Cohort.csv")
daps <-dapss[, -1]
```

```{r, include=FALSE}
#### only run this once so that the data sets don't get mixed up ####

names(daps) #find the right columns
pollution <- daps[,c(1:3, 27:40, 56:61)] #subset daps to only pollution data for simplification
names(pollution) #check that we have the right names

id <- daps$id
visit <- daps$visit
season <- daps$season

```

Part of what we're interested in is estimating the mean bias and mean variance. If we assume that the "true" value is that from the wearable monitor, how do we estimate the mean bias? What about the mean variance?

Variance is defined as:
\[\textrm{Var}[ \widehat{x} ]=E[(\widehat{x} -E[\widehat{x}])^{2} ] \]

\[ \textrm{Var}[ \widehat{x}]=E[\widehat{x}^2]-E[\widehat{x}]^2 \]

Bias is defined as follows:

\[ \textrm{Bias}[ \widehat{x}]=E[\widehat{x}-x]=E[\widehat{x}]-x \]

```{r, echo=FALSE}

##### PM10 #####
wearable <- exp(daps$log_vest_pm10)
stationary <- exp(daps$log_msp_pm10)
cdphe_mean <- (daps$CDPHE_PM10_mean)
cdphe_max <- exp(daps$log_CDPHE_PM10_max)


PM10 <- as.data.frame(cbind(id, visit, season, wearable, stationary, cdphe_mean, cdphe_max))

pm10 <- melt(PM10, id.vars = c("id", "visit", "season"), measure.vars = c( "wearable", "stationary","cdphe_mean","cdphe_max"), variable.name = "type", value.name = "PM10",
  factorsAsStrings = TRUE)

names(pm10) <- c("id", "visit", "season", "type", "PM10")


#Understanding the mean bias and variance between the wearable (our "truth") and the two other possible monitors. 

#Mean Bias of the CDPHE MEAN 
bias_cdphe <-mean(mean(cdphe_mean, na.rm=TRUE) - wearable, na.rm=TRUE)

#Mean Variance of the CDPHE MEAN:

  #Mean of the Sample Squared:
ss <- mean(cdphe_mean*cdphe_mean, na.rm=TRUE)

  #Mean of the Squared Sample:
ess <-(mean(cdphe_mean, na.rm=TRUE)*mean(cdphe_mean, na.rm=TRUE))

mvar_cdphe <- ss-ess

var_cdphe <- var(cdphe_mean, na.rm=TRUE)

#Mean Bias of the Stationary Monitor
bias_stat <-mean(mean(stationary, na.rm=TRUE)-wearable, na.rm=TRUE)
#Mean Variance of the Stationary Monitor
  #Mean of the Sample Squared:
ss <- mean(stationary*stationary, na.rm=TRUE)

  #Mean of the Squared Sample:
ess <-(mean(stationary, na.rm=TRUE)*mean(stationary, na.rm=TRUE))

mvar_stat <- ss-ess

var_stat <- var(stationary, na.rm=TRUE)

#then, we repeat the means of the stationary and the cdphe across the length of the wearable values
mean_stat <- rep((mean(stationary, na.rm=TRUE)), length(wearable))
mean_cdphe <- rep((mean(cdphe_mean, na.rm=TRUE)), length(wearable))

#and create a new data frame comparing the stationary/cdphe means to each value of the wearable. 
dat2 <- as.data.frame(cbind(wearable, mean_stat, mean_cdphe))

#make two vectors of the bias, the mean variance, and the variance of the cdphe and the stationary respectively to print
CDPHE_VAL <- c(bias_cdphe, mvar_cdphe, var_cdphe)
STAT_VAL <- c(bias_stat, mvar_stat, var_stat)
```
### CDPHE PM10: Bias, Mean Variance, Variance
```{r, echo=FALSE}
CDPHE_VAL
```
### Stationary PM10: Bias, Mean Variance, Variance
```{r, echo=FALSE}
STAT_VAL
```






```{r, echo=FALSE}
#### NO2 #####
wearable <- exp(daps$log_vest_NO2)
stationary <- exp(daps$log_stat_NO2)
cdphe_mean <- exp(daps$log_CDPHE_NO2_mean)
cdphe_max <- exp(daps$log_CDPHE_NO2_max)
  
NO2 <- as.data.frame(cbind(id, visit, season, wearable, stationary, cdphe_mean, cdphe_max))

no2 <- melt(NO2, id.vars = c("id", "visit", "season"), measure.vars = c( "wearable", "stationary","cdphe_mean","cdphe_max"),
  variable.name = "type", value.name = "NO2",
  factorsAsStrings = TRUE)

names(no2) <- c("id", "visit", "season", "type", "NO2")

#Mean Bias of the CDPHE MEAN 
bias_cdphe <- mean(mean(cdphe_mean, na.rm=TRUE) - wearable, na.rm=TRUE)



#Mean Variance of the CDPHE MEAN:

  #Mean of the Sample Squared:
ss <- mean(cdphe_mean*cdphe_mean, na.rm=TRUE)

  #Mean of the Squared Sample:
ess <-(mean(cdphe_mean, na.rm=TRUE)*mean(cdphe_mean, na.rm=TRUE))

mvar_cdphe <- ss-ess

var_cdphe <- var(cdphe_mean, na.rm=TRUE)

#Mean Bias of the Stationary Monitor
bias_stat <- mean(mean(stationary, na.rm=TRUE)-wearable, na.rm=TRUE)
#Mean Variance of the Stationary Monitor
  #Mean of the Sample Squared:
ss <- mean(stationary*stationary, na.rm=TRUE)

  #Mean of the Squared Sample:
ess <-(mean(stationary, na.rm=TRUE)*mean(stationary, na.rm=TRUE))

mvar_stat <-ss-ess

var_stat <- var(stationary, na.rm=TRUE)


#then, we repeat the means of the stationary and the cdphe across the length of the wearable values
mean_stat <- rep((mean(stationary, na.rm=TRUE)), length(wearable))
mean_cdphe <- rep((mean(cdphe_mean, na.rm=TRUE)), length(wearable))

#and create a new data frame comparing the stationary/cdphe means to each value of the wearable. 
dat2 <- as.data.frame(cbind(wearable, mean_stat, mean_cdphe))


#make two vectors of the bias, the mean variance, and the variance of the cdphe and the stationary respectively to print
CDPHE_VAL <- c(bias_cdphe, mvar_cdphe, var_cdphe)
STAT_VAL <- c(bias_stat, mvar_stat, var_stat)
```

### CDPHE NO2: Bias, Mean Variance, Variance
```{r, echo=FALSE}
CDPHE_VAL
```
### Stationary NO2: Bias, Mean Variance, Variance
```{r, echo=FALSE}
STAT_VAL
```


```{r, echo=FALSE}
#### O3 #####
wearable <- daps$vest_O3
cdphe_mean <- daps$CDPHE_O3_mean
cdphe_max <- daps$CDPHE_O3_max

O3 <- as.data.frame(cbind(id, visit, season, wearable, cdphe_mean, cdphe_max))

o3 <- melt(O3, id.vars = c("id", "visit", "season"), measure.vars = c( "wearable", "cdphe_mean", "cdphe_max"),
  variable.name = "type", value.name = "O3",
  factorsAsStrings = TRUE)

names(o3) <- c("id", "visit", "season", "type", "O3")

#Mean Bias of the CDPHE MEAN 
bias_cdphe <- mean(mean(cdphe_mean, na.rm=TRUE) - wearable, na.rm=TRUE)

#Mean Variance of the CDPHE MEAN:

  #Mean of the Sample Squared:
ss <- mean(cdphe_mean*cdphe_mean, na.rm=TRUE)

  #Mean of the Squared Sample:
ess <-(mean(cdphe_mean, na.rm=TRUE)*mean(cdphe_mean, na.rm=TRUE))

#mean variance
mvar_cdphe <- ss-ess

#actual variance
var_cdphe <- var(cdphe_mean, na.rm=TRUE)

#make a vector to compare against
mean_cdphe <- rep((mean(cdphe_mean, na.rm=TRUE)), length(wearable))

dat2 <- as.data.frame(cbind(wearable, mean_cdphe))
#make a vector to show the bias, mean variance, and variance next to each other
CDPHE_VAL <- c(bias_cdphe, mvar_cdphe, var_cdphe)

```

### CDPHE O3: Bias, Mean Variance, Variance
```{r, echo=FALSE}
CDPHE_VAL
```


#   Evaluating differences between monitors using mixed models. 

## PM10, MEASUREMENT DIFFERENCES

Evaluation using a mixed model with random intercept for each id. For the first model reference categories are the wearable monitor and summer. We see a slight significant difference between the wearable monitor and the stationary monitor and a significant difference between the wearable monitor and the CDPHE monitor. Additionally, the fall has significantly higher PM10 values than the summer. In the second model, we see again the slight difference between the wearable and stationary monitor but also a slight difference between the stationary and the CDPHE. 


```{r, echo=FALSE}

pm10_3 <- subset(pm10, type != "cdphe_max")

pm10_3$season <- as.factor(pm10_3$season)
pm10_2 <- pm10_3

pm10_5 <- pm10_3


#summary(pm10monit2)



pm10_2<- within(pm10_2, type <- relevel(type, ref = "wearable"))


pm10_5<- within(pm10_5, type <- relevel(type, ref = "stationary"))

#with the personal monitor as the reference category
pm10monit <- lmer(PM10 ~ type + season + (1|id), data = pm10_2 )
pm10monit2 <- lmer(PM10 ~ type + season + (1|id), data = pm10_5 )


#summary(pm10monit)
#summary(pm10monit2)

class(pm10monit) <- "lmerMod"
class(pm10monit2) <- "lmerMod"

```

```{r, echo=FALSE}
stargazer(pm10monit, type="text", style="all", ci=T, ci.level=0.95, summary=FALSE, digits=4, star.cutoffs=c(0.05, 0.01, 0.001), title="PM10 Monitor Differences,ref = wearable", covariate.labels=c("Stationary", "CDPHE", "Fall", "Winter", "Spring", "Intercept"))
```

```{r, echo=FALSE}

stargazer(pm10monit2, type="text", style="all", ci=T, ci.level=0.95, summary=FALSE, digits=4, star.cutoffs=c(0.05, 0.01, 0.001), title="PM10 Monitor Differences, ref=stationary", covariate.labels=c("Wearable", "CDPHE", "Fall", "Winter", "Spring", "Intercept"))

```


```{r, echo=FALSE}

levels(pm10_5$type )[levels(pm10_5$type )=="cdphe_mean"] <- "CDPHE"
#BOXPLOT
p <- ggplot(pm10_5, aes(type, PM10))
p + geom_boxplot()+
  xlab("Monitor")+
  ylab("PM10")+ 
  labs(title = "Measured Pollutant by Monitor Type, w/outlier", color ="Type")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  theme(plot.title=element_text(face="bold", size=15)) +theme(text = element_text(size = 15, color= "black"))

#SMOOTHED PLOTS OVER TIME
b <- ggplot(pm10_5, aes(as.numeric(season), PM10, color=type, shape=type))
b + geom_point()+ xlab("Season") + ylab(expression( PM[10]))+ labs(title = "Measured Pollutant by Monitor Type, w/outlier", color ="Type", shape ="Type")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(plot.title=element_text( face="bold", size=15)) + geom_smooth() + scale_x_continuous(labels=c("1" = "Summer", "2" = "Fall", "3"= "Winter", "4"= "Spring"))+ scale_color_manual(values = c("wearable" = "blue", "stationary" = "limegreen", "CDPHE" = "red")) +theme(text = element_text(size = 15, color= "black"))

#WITH THE OUTLIERS REMOVED
pm10_3 <- subset(pm10_5, pm10_5$PM10 < 200)
b <- ggplot(pm10_3, aes(as.numeric(season), PM10, color=type, shape=type))
b + geom_point()+ xlab("Season") + ylab(expression( PM[10]))+ labs(title = "Measured Pollutant by Monitor Type, w/o outlier", color ="Type", shape ="Type")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+ scale_color_manual(values = c("wearable" = "blue", "stationary" = "limegreen", "CDPHE" = "red"))+theme(plot.title=element_text( face="bold", size=15)) + geom_smooth() + scale_x_continuous(labels=c("1" = "Summer", "2" = "Fall", "3"= "Winter", "4"= "Spring"))+theme(text = element_text(size = 15, color= "black")) 





```





## NO2 MEASUREMENT DIFFERENCES
Evaluation using a mixed model with random intercept for each id. The reference categories are the wearable monitor and summer for the first model. We see a significant difference between CDPHE measures and the wearable monitor. In the second model, the reference categories are the stationary monitor and summer. We see that there is also a significant difference between the CDPHE monitor and the stationary monitor. The stationary monitor and the wearable monitor are not significantly different from each other. 


```{r, echo=FALSE}
#### NO2 #####


no2_3 <- subset(no2, type != "cdphe_max")

no2_3$season <- as.factor(no2_3$season)
no2_2 <- no2_3

no2_5 <- no2_3


#summary(pm10monit2)



no2_2<- within(no2_2, type <- relevel(type, ref = "wearable"))


no2_5<- within(no2_5, type <- relevel(type, ref = "stationary"))

#with the personal monitor as the reference category
no2monit <- lmer(NO2 ~ type + season + (1|id), data = no2_2 )
no2monit2 <- lmer(NO2 ~ type + season + (1|id), data = no2_5 )


#summary(no2monit)
#summary(no2monit2)

class(no2monit) <- "lmerMod"
class(no2monit2) <- "lmerMod"

```

```{r, echo=FALSE}
stargazer(no2monit, type="text", style="all", ci=T, ci.level=0.95, summary=FALSE, digits=4, star.cutoffs=c(0.05, 0.01, 0.001), title="NO2 Monitor Differences,ref = wearable", covariate.labels=c("Stationary", "CDPHE", "Fall", "Winter", "Spring", "Intercept"))
```


```{r, echo=FALSE}
stargazer(no2monit2, type="text", style="all", ci=T, ci.level=0.95, summary=FALSE, digits=4, star.cutoffs=c(0.05, 0.01, 0.001), title="NO2 Monitor Differences, ref=stationary", covariate.labels=c("Wearable", "CDPHE", "Fall", "Winter", "Spring", "Intercept"))

```

```{r, echo=FALSE}
levels(no2_5$type )[levels(no2_5$type )=="cdphe_mean"] <- "CDPHE"


p <- ggplot(no2_5, aes(type, NO2))
p + geom_boxplot()+xlab("Monitor")+ ylab("NO2")+ labs(title = "Measured Pollutant by Monitor Type", color ="Monitor")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(plot.title=element_text( face="bold", size=15)) +theme(text = element_text(size = 15, color= "black"))

b <- ggplot(no2_5, aes(as.numeric(season), NO2, color=type, shape =type))
b + geom_point()+ xlab("Season") + ylab("NO2")+ labs(title = "Measured Pollutant by Monitor Type, w/outlier", color ="Monitor", shape ="Monitor")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(plot.title=element_text(face="bold", size=15)) + geom_smooth()+ scale_color_manual(values = c("wearable" = "blue", "stationary" = "limegreen", "CDPHE" = "red")) + scale_x_continuous(labels=c("1" = "Summer", "2" = "Fall", "3"= "Winter", "4"= "Spring"))+theme(text = element_text(size = 15, color= "black"))

b <- ggplot(no2_5, aes(as.numeric(season), NO2, color=type, shape =type))
b + geom_point()+ xlab("Season") + ylab("NO2")+ labs(title = "Measured Pollutant by Monitor Type, w/o outlier", color ="Monitor", shape ="Monitor")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(plot.title=element_text(face="bold", size=15)) + scale_color_manual(values = c("wearable" = "blue", "stationary" = "limegreen", "CDPHE" = "red"))+ geom_smooth() + scale_x_continuous(labels=c("1" = "Summer", "2" = "Fall", "3"= "Winter", "4"= "Spring"))+theme(text = element_text(size = 15, color= "black"))



```



# OZONE MEASUREMENT DIFFERENCES

We see an unsurprising and substantial difference between the CDPHE mean value and the wearable monitor.

```{r, echo=FALSE}
library(plyr)
levels(o3$type)[levels(o3$type)=="cdphe_mean"] <- "CDPHE"
o3_2 <- subset(o3, type != "cdphe_max")

#### O3 #####
o3monit <- lm(O3 ~ as.factor(type), data = o3_2 )
#summary(o3monit)


#the t-tests should give similar results
#t.test(O3$vest, O3$cdphe_mean, paired=TRUE, conf.level=0.95)


p <- ggplot(o3_2, aes(type, O3))
p + geom_boxplot() +xlab("Monitor")+ ylab("Ozone")+ labs(title = "Measured Pollutant by Monitor Type", color ="Type")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(plot.title=element_text( face="bold", size=15)) +theme(text = element_text(size = 15, color= "black"))

stargazer(o3monit, type="text", style="all", ci=T, ci.level=0.95, summary=FALSE, digits=4, star.cutoffs=c(0.05, 0.01, 0.001), title="Ozone Monitor Differences", covariate.labels=c("CDPHE", "Intercept (Wearable)"))


```

