---
title: "DAPS_Data Management"
author: "Melissa Lowe"
date: "5/23/19"
output: html_document
---

This was rerun with new carbon data to produce new clean datasets. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries
```{r}
library(readxl)
library(utils)
library(reshape2)
library(plyr)
```

Original Data Files

```{r}
setwd("~/Desktop/ReproducibleItemsforDAPS") 

name.gluc <- c("visit", "num", "id", "sample", "pm_wt", "endo_ml", "gluc_ml", "endo_m3", "gluc_m3", "note")
name.demo <- c("id", "visit", "season", "cohort", "X__1", "age", "gender", "race_afam", "X__2","X__3","X__4","X__5","X__6","X__7")

end_gluc <- read_excel("endoglucanfilters_09272018.xlsx", col_names = name.gluc, skip=2)
demographics <- read_excel("DAPS_Primary_DataSet_180902.xlsx", col_names=name.demo, sheet = "Demographic Predictors", skip=1)
cont_outcomes<- read_excel("DAPS_Primary_DataSet_180902.xlsx", sheet = "Continuous Outcome Variables")
cont_exp_var <- read_excel("DAPS_Primary_DataSet_180902.xlsx", sheet = "Continuous Exposure Variables")
cat_exp_var <- read_excel("DAPS_Primary_DataSet_180902.xlsx", sheet = "Zero-Heavy & Categ. Exposures")
cat_outcomes<- read_excel("DAPS_Primary_DataSet_180902.xlsx", sheet = "Categorical Outcomes")
exacerbation_count <- read_excel("DAPS_Primary_DataSet_180902.xlsx", sheet = "ExacerbationsSinceLastVisit")
newcarbon <- read_excel("DAPS_CarbonData_112818.xlsx")
```
Making sure that the data is in the right format
```{r}
end_gluc$visit <- as.numeric(end_gluc$visit)
end_gluc$endo_ml <- as.numeric(end_gluc$endo_ml)
end_gluc$endo_m3 <- as.numeric(end_gluc$endo_m3)
end_gluc$gluc_m3 <- as.numeric(end_gluc$gluc_m3)


unique(end_gluc$id)
```
Merging

```{r}
merge1 <- merge(demographics, cont_outcomes, by = c("id","visit","season","cohort", "X__1"), all=TRUE)
merge2 <- merge(merge1, cont_exp_var, by = c("id","visit","season","cohort", "X__1"), all=TRUE)
merge3 <- merge(merge2, cat_outcomes, by = c("id","visit","season","cohort", "X__1"), all=TRUE)
merge4 <- merge(merge3, cat_exp_var, by = c("id","visit","season","cohort", "X__1"), all=TRUE)
merge5 <- merge(merge4, end_gluc, by = c("id", "visit"), all.x=TRUE)

merge6 <- merge(merge5, newcarbon, by = c("id", "visit", "X__1", "X__2","X__3", "X__4"), all.x=TRUE)

#get rid of excess columns and old data
merge6 <- subset(merge6, select = -c(X__1,X__2,X__3,X__4, X__5, X__6, X__7, vest_ETS_adj0,vest_BrC_adj0, vest_BC_adj0, msp_ETS_adj0, msp_BrC_adj0, msp_BC_adj0))

#change column names

col_merge6 <-c("id","visit", "season", "cohort","age","gender","race_afam", "FeNO_avg","log_FeNO_avg","FEV1_pp_GLI","FEV1_Zscore_GLI", "FVC_pp_GLI", "FVC_Zscore_GLI","FEV1_FVC", "FEV1_FVC_pp_GLI","FEV1_FVC_Zscore_GLI" ,"FEF2575_pp_GLI","FEF2575_Zscore_GLI","cbc_wbc","log_cbc_wbc" ,
               "cbc_abs_eos","cbc_abs_neut","log_cbc_abs_neut","cbc_percent_eos",
               "cbc_percent_neut","log_cbc_percent_neut","log_vest_pm10","log_vest_ETS","log_vest_BrC", "log_msp_pm10","log_msp_BC","log_vest_NO2",
               "log_stat_NO2","vest_O3","CDPHE_PM10_mean","log_CDPHE_PM10_max",
               "log_CDPHE_NO2_mean","log_CDPHE_NO2_max" ,"CDPHE_O3_mean","CDPHE_O3_max","CASI_Day","CASI_Night", "CASI_Function","CASI_Medication","CASI_Exacerbation","CASI_Exac_5a","CASI_Exac_5b","CASI_Total_NoFxn","CASI_Total_wFxn",
               "CACT_ACT_Cat","CACT_ACT_cont", "Burst_2months","Hospital_2months", "DrVisit_2months","ExacerbationAtVisit","nicalert", 
               "num" ,"sample","pm_wt","endo_ml", "gluc_ml","endo_m3","gluc_m3", "note","vest_ETS","vest_BrC", "vest_BC","vest_ETS-BrC_MDL","vest_BC_MDL", "MSP_ETS","MSP_BrC","MSP_BC","MSP_ETS_BrC_MDL", "MSP_BC_MDL")
colnames(merge6)  <- col_merge6   

merge6 <- subset(merge6, select = -c(log_vest_ETS,log_vest_BrC,log_msp_BC))



```






Dealing with Exacerbation Data:

ID is shared, cohort is shared, exposure visit matches with normal visit number, exposure season matches with normal season value, outcome visit and outcome season are different.

So, to figure out how to merge them, we'll do all of the final modifications of variable types last.

```{r}
#first, drop unnecessary columns.
merge6a <- subset(exacerbation_count, select = -c(outcome_visit,outcome_season))

#create a column marking if there was ever an exacerbation:

merge6a$exacsum <- merge6a$Burst_SinceLastVisit+ merge6a$Hospital_SinceLastVisit + merge6a$DrVisit_SinceLastVisit

merge6a$ynexac <- ifelse(merge6a$exacsum > 0, 1,0)

name6a <- c("id", "cohort", "visit", "season", "Burst", "Hosp", "Dr", "exacsum", "ynexac")

colnames(merge6a)  <- name6a   

merge6b <- merge(merge6, merge6a, by = c("id","visit","season","cohort"), all=TRUE)

merge6b <- subset(merge6b, select= -c(Burst_2months, Hospital_2months, DrVisit_2months, ExacerbationAtVisit))


```








```{r}

#Modifying Labels to make interpretation clearer.

merge6b$gender[is.na(merge6b$gender)] <- "Missing"
merge6b$race_afam[is.na(merge6b$race_afam)] <- "Missing"
merge6b$race_afam <- ifelse(merge6b$race_afam ==1, "Yes", merge6b$race_afam)
merge6b$race_afam <- ifelse(merge6b$race_afam == 0, "No", merge6b$race_afam)

merge6b$age <- as.numeric(merge6b$age)
merge6b$ageCat <- cut(merge6b$age,breaks=c(7,11,17),labels=c("8-11","12-17"))
merge6b$CASI_Total_wFxnCat <- cut(merge6b$CASI_Total_wFxn,breaks=c(3,4,20),labels=c("low","high"))
merge6b$season_2 <- ifelse(merge6b$season == 4,3,merge6b$season)

merge6b$season <- as.factor(merge6b$season)
merge6b$season <- revalue(merge6b$season, c("1"="summer","2"="fall", "3"= "winter", "4"="spring"))


```



Create the final datasets for export.
```{r}
merge7 <- subset(merge6b, cohort== 2)


mergea <- subset(merge6b, cohort == 1)

```


Export csv files. The names of the output files to be used for further DAPS analysis are:

**DAPS_AllCohort.csv**
**DAPS_2Cohort.csv**
**DAPS_1Cohort.csv**


```{r}

setwd("~/Desktop/ReproducibleItemsforDAPS") 
write.csv(merge6b, file = "DAPS_AllCohort.csv", na= "NA") #all subjects
write.csv(merge7, file = "DAPS_2Cohort.csv", na = "NA") #just cohort 2
write.csv(mergea, file = "DAPS_1Cohort.csv", na = "NA") #just cohort 1

```
